This dissertation would not have been possible without the help of many people including my family, my friends, and everyone at Stanford who makes research possible. 
First, I would like thank my reading committee, Professor Dorsa Sadigh and Professor Marco Pavone, for providing valuable feedback on my research, defense, and dissertation. Their perspective and inquiry helped hone my research direction, and both have provided valuable advice and opportunities outside of my dissertation work. Then, I would like to thank my advisor Professor Mykel Kochenderfer for mentoring me in research and in life. Mykel helped me discover the wonders of machine learning and decision making, taught me how to communicate more clearly and succinctly, and demonstrated how to live a life full of passion, gratitude, and kindness. I am incredibly grateful for everything I have learned from Mykel, and look forward to our continued collaborations as co-workers and friends. 

My first three years at Stanford were funded by the Lillie Family Stanford Graduate Fellowship. Without that fellowship I may not have attended Stanford, so I thank the Aeronautics and Astronautics department for nominating me and the Lillie Family for providing the funds. I want to thank the Stanford Center for AI safety for funding my dissertation research, and specifically NVIDIA for providing the funding and research direction. Thank you to Michael Cox and Ahmed Nassar at NVIDIA for always providing feedback on my work and supporting the safety center. Thank you to NASA Ames and Dr. Ritchie Lee for supporting my summer internship at NASA Ames and collaborating on several projects. Lastly, I would like to thank the NASA University Leadership Initiative for supporting my last quarter of research, and Professor Marco Pavone for providing the opportunity.

I have had the opportunity to collaborate with many amazing friends and researchers. I would first like to thank Dr. Ashish Goel, Dr. Paul Tarantino and especially Dr. Siddarth Krishnamoorthy for giving me valuable research experience and advice as I navigated my first several years of graduate school. Thank you to Dr. Jeremy Morton, Professor Freddie Witherden, Mark Koren, Xiaobai Ma, Peter Du, and Professor Katherine Driggs-Campbell, for collaborating on my first research projects as a member of SISL. Thank you to Dr. Ritchie Lee for developing many of the ideas that I built on for my thesis, and for being a valuable research mentor. Thank you to Robert Moss, Dr. Alex Koufos, Dr. Maxime Bouton, Dr. Di Wu, and all of the other SISL members who gave me advice and feedback on my research. Lastly, thank you to Sydney Katz and Chris Strong for helping me develop my next stage of research projects. 

Thank you to my roommate Stephen for being a supportive and easy going living partner. Thank you to Krishna, Jasper, Jenny, John, and all of the other friends I have met (and climbed with) at Stanford, as well as my friends from Harvey Mudd and back home in Oregon. Thank you to my girlfriend Emma. She provided an endless source of advice, support, and love during my time in graduate school. Thank you for always having my back, cheering me on, and always being willing to share a laugh or go on an adventure. 

Lastly, thank you to my family, to whom I truly owe everything. Thank you for teaching me to be the person I am today and providing me all the support I have needed to pursue my dreams. Thank you to my sister Amie, who can always make me feel better when I am in a bad spot, and whose enthusiasm about the work we do always leads to great conversations. Thank you to my Mom and Dad who have always encouraged me to think for myself and work hard for the things that are important. I love you all.
